---
title: 'Name Method: tables for the manuscript'
author: "Cal Chengqi Fang"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    keep_tex: TRUE
classoption: landscape
header-includes:
  - \usepackage{float}
knit: >
  (function(inputFile, encoding){
    rmarkdown::render(inputFile, encoding=encoding, output_file="results/tables/manuscriptTables.pdf")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      message=FALSE,
                      warning=FALSE)

require(tidyverse)
require(data.table)

require(knitr)      # For generating nice and organized results
require(kableExtra) # For generating nice and organized results
```

```{r data}
# Load the formatted matrices
load(file="data/interm/formatted.Rdata")

# Collapse matrices back to summary dataframe
cntrmat_triplet <- as(cntrmat, "TsparseMatrix")
cntr <- data.frame(area = rownames(cntrmat)[cntrmat_triplet@i + 1],
                   surname = colnames(cntrmat)[cntrmat_triplet@j + 1],
                   freq = cntrmat_triplet@x)

ntvmat_triplet <- as(ntvmat, "TsparseMatrix")
ntv <- data.frame(area = rownames(ntvmat)[ntvmat_triplet@i + 1],
                  surname = colnames(ntvmat)[ntvmat_triplet@j + 1],
                  freq = ntvmat_triplet@x)

# Load the reportable lists
load(file="data/interm/resultRep.Rdata")

# Load the validation summary with the Census list
load("data/interm/validateCensus.Rdata")

# Load the validation summary through cross-validation
load("data/interm/validateCV.Rdata")
```

```{r prep}
# Create a lookup table
codeLookup <- data.frame(code = c("AF", "BIA", "BM", "BT", "CB", 
                                  "CE", 
                                  "GCA", 
                                  "JA", "LA", "MG", "MV", 
                                  "MY", "NP", "Other", 
                                  "RP", "SAS", 
                                  "SN", "TH", 
                                  "VM", 
                                  "XK"),
                         name = c("Afghanistan", "Brunei, Indonesia", "Myanmar", "Bhutan", "Cambodia",
                                  "Sri Lanka", 
                                  "Peoples Republic of China, Hong Kong, Macau (Macao), Taiwan", 
                                  "Japan, Southern Ryukyu Islands", "Laos", "Mongolia", "Maldives",
                                  "Malaysia", "Nepal", "All the other countries and regions", 
                                  "Philippines", "Bangladesh, India, Pakistan, Sikkim, British India", 
                                  "Singapore", "Thailand", 
                                  "Vietnam, Democratic Republic of Vietnam, Republic of Vietnam", 
                                  "Korea, Democratic People’s Republic of Korea, Republic of Korea"))

# Write a function to replace area codes with names and sort the table
replace_sort <- function(data, code_col, lookup=codeLookup) {
  # Merge the data with the lookup table
  data <- merge(data, lookup, 
                by.x=code_col, by.y="code", all.x=TRUE)
  
  # Remove the original column
  data[[code_col]] <- NULL
  
  # Extract the Other row
  special_row <- data %>% 
    filter(name == "All the other countries and regions")
  
  # Remove the special row from the data
  data <- data %>% 
    filter(name != "All the other countries and regions")
  
  # Sort the remaining data
  data <- data %>% 
    arrange(name)
  
  # Bind the special row at the bottom
  data <- bind_rows(data, special_row)
  
  return(data)
}
```
In this document, I made some tables for our manuscript to illustrate our analysis process and results.

\newpage
# Table 1. Illustrative surnames and corresponding conditional probabilities
```{r tab1}
nameSample <- str_pad(c("HUANG", "CHANG", "KIM", "SUZUKI", "NGUYEN", "INTHAVONG", "PENAFLOR", "SAMEER", "PATAN"),
                      12, side="right")
tab1 <- probList %>% 
  filter(surname %in% str_pad(nameSample, 12, side="right")) %>% 
  mutate(surname = factor(surname, levels=nameSample)) %>% 
  arrange(surname) %>% 
  select(-surname) %>% 
  mutate(across(.cols=everything(), .fns=~round(.x, 4)*100)) %>% 
  t() %>% 
  as_tibble() %>% 
  mutate(endCol = colnames(probList)[1:20])
colnames(tab1) <- c(nameSample, "area")

# Replace area codes with actual names
tab1 <- replace_sort(tab1, "area") %>% 
  select(name, nameSample)

tab1 %>% 
  kbl(col.names=c("Country/Area of origin", nameSample),
      align="l", booktabs=TRUE) %>% 
  kable_styling(latex_options = c("hold_position")) %>%
  column_spec(1, width="5cm") %>% 
  add_header_above(header=c(" "=1, "Illustrative surnames[note]"=9)) %>% 
  add_footnote("The probabilities in each column do not add up to 1 because of rounding error.",
               notation="symbol",
               threeparttable=TRUE)
```

\newpage
# Table 2. Predictive surnames and related record numbers
```{r tab2}
tab2Area <- function(areaCode){
  # Identify surnames associated with the corresponding areaCode
  tab2Names <- determList %>% 
    filter(area == areaCode)
  
  # Make column 1: Country/Area
  col1 <- areaCode
  # Make column 2: Number of highly predictive surnames
  col2 <- nrow(tab2Names)
  # Make column 3: Number of records in the US-born with these highly predictive surnames
  tab2Ntv <- ntv %>% 
    filter(surname %in% tab2Names$surname) 
  col3 <- sum(tab2Ntv$freq)
  # Make column 4: Number of records in the foreign-born file with these highly predictive surnames
  tab2Cntr <- cntr %>% 
    filter(surname %in% tab2Names$surname) 
  col4 <- sum(tab2Cntr$freq)
  
  # Construct the result
  result <- tibble(area = col1,
                   nameNum = col2,
                   ntvTotal = col3,
                   cntrTotal = col4)
}

tab2 <- lapply(codeLookup$code, tab2Area) %>% 
  data.table::rbindlist()

# Replace area codes with actual names
tab2 <- replace_sort(tab2, "area") %>% 
  select(name, nameNum, ntvTotal, cntrTotal) %>% 
  filter(name != "All the other countries and regions")

tab2 %>% 
  kbl(col.names=c("Country/Area of origin", "Number of predictive surnames[note]", 
                  "Number of US-born records with corresponding predictive surnames",
                  "Number of foreign-born records with corresponding predictive surnames"),
      format.args=c(big.mark=","), align="l", booktabs=TRUE) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  column_spec(1, width="5cm") %>% 
  column_spec(2:4, width="3cm") %>% 
  add_footnote("The numbers do not add up to 22,621 because there are surnames that are equally predictive of more than one country/area.",
               notation="symbol",
               threeparttable=TRUE)
```

\newpage
# Table 3. Validation result with ACS data
```{r tab3}
countrySummary <- predictSummary %>% 
  filter(maxCntr != "Whole List") %>% 
  rbind(c("MV", 0, NA, NA, NA))

# Replace area codes with actual names
countrySummary <- replace_sort(countrySummary, "maxCntr") %>% 
  select(name, nameCount, PPV, Sensitivity, Specificity) %>% 
  filter(name != "All the other countries and regions")

tab3 <- predictSummary %>% 
  filter(maxCntr == "Whole List") %>% 
  rename(name = maxCntr) %>% 
  rbind(countrySummary) %>% 
  mutate(across(.cols=PPV:Specificity, .fns=~round(as.numeric(.x), 4)*100))

tab3 %>% 
  kbl(col.names=c(" ", "Number of surnames present in both lists[note]", 
                  "PPV (%)", "Sensitivity (%)", "Specificity (%)"),
      align="l", booktabs=TRUE) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  column_spec(1, width="5cm") %>% 
  column_spec(2:5, width="3cm") %>% 
  add_footnote("The numbers of surnames of countries/areas do not add up to 3,370 because there are surnames that are equally predictive of more than one country/area.",
               notation="symbol",
               threeparttable=TRUE) %>% 
  pack_rows("", 2, 20)
```

\newpage
# Table 4. Validation result of cross validation
```{r tab4}
tab4 <- cvResultsSummary %>% 
  select(area:`Pos Pred Value`) 

# Replace area codes with actual names
tab4 <- replace_sort(tab4, "area") %>% 
  select(name, `Pos Pred Value`, Sensitivity, Specificity) %>% 
  filter(name != "All the other countries and regions") %>% 
  mutate(across(.cols=`Pos Pred Value`:Specificity, .fns=~round(as.numeric(.x), 4)*100))

tab4 %>% 
  kbl(col.names=c(" ", "PPV (%)", "Sensitivity (%)", "Specificity (%)"),
      align="l", booktabs=TRUE) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  column_spec(1, width="5cm") %>% 
  column_spec(2:4, width="3cm")
```

\newpage
# Sample Table. Numbers of foreign-born records from corresponding country/area - Leftout
```{r sample tab}
# Create sample tab to show the foreign-born frequencies
smpTab <- cntr %>% 
  group_by(area) %>% 
  summarise(obs = sum(freq)) 

# Replace area codes with actual names
smpTab <- replace_sort(smpTab, "area") %>% 
  select(name, obs)

smpTab %>% 
  kbl(col.names=c("Country/Area of birth", "Number of foreign-born records"),
      format.args=c(big.mark=","), align="l", booktabs=TRUE) %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  column_spec(1:2, width="5cm")
```
